---
title: Visualisation of each Nigeria state contribution \n to the IGR of the Country
  in the first Quater of 2021
author: "Akinola Samson"
date: "1/6/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
Data used was downloaded from the National bureau of statistics website,the  appropriate Libraries were loaded and the data was imported,removing unnecessary information with the aid of skip and n_max during the process
```{r Loading, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
Half_Year_2021 <- read_excel("~/Half_Year_2021.xlsx", 
    skip = 1, n_max = 37)
```

After importing the data and loading of necesary packages,the data was processed in preparation for analysis by creating a new column names per_tax wich shows the percentage contribution of each state by dividing the grand total of each state by sum of the grand total for all the states
```{r}
df<- Half_Year_2021 %>%
  mutate(per_tax= Half_Year_2021$`Grand Total`/sum(Half_Year_2021$`Grand Total`))
```
A new data list was created to show the geopolitical zone of each states and added to the data frame


```{r }
zone<- c("South East", "North East","South South",
                                     "South East","North East","South South",
                                     "North Central","North East","South South",
                                     "South South","South East","South South",
                                     "South West","South East","North East",
                                     "South East","North West","North West",
                                     "North West","North West","North West",
                                     "North Central","North Central","South West",
                                     "North Central","North Central","South West",
                                     "South West","South West","South West",
                                     "North Central","South South","North West",
                                     "North East","North East","North West",
                                     "North Central")
df$Zone<- zone
```
A scattered plot was used to visualize the contribution of each state by mapping per_tax to x axis and state(ordered by per_tax) to the y axis.
the scale on the x axis was changed to logarithm to properly accomodate the whole states including the outlier (Lagos) and scaled by percentaged.Colour was mapped by geopolitical zone with label positioned by the right side of the map

```{r PlotE}
p<- ggplot(data = df, mapping = aes(x= per_tax, y= reorder(State, per_tax)))
p+ geom_point(mapping = aes(color=Zone))+ scale_x_log10(labels= scales::percent)+
  labs(x= "Percentage", y= "",
       title = "Percentage contribution of Nigeria states to the Country's IGR",
       subtitle = "First Quater of 2021",
       caption = "Data Source: National Bureau of Statistic",
       color= "Geopolitical Zones")
```

A line showing 1% mark was added to the plot to demarcate states contributing less that 1% of the IGR
```{r}
p+ geom_point(mapping = aes(color=Zone))+ scale_x_log10(labels= scales::percent)+
  labs(x= "Percentage", y= "",
       title = "Percentage contribution of Nigeria states to the Country's IGR",
       subtitle = "First Quater of 2021",
       caption = "Data Source: National Bureau of Statistic",
       color= "Geopolitical Zones")+
  geom_vline(xintercept = 0.01, color="red", size= 0.5, alpha=0.4)+
  annotate(geom = "text", x= 0.009, y=25, label= "Line showing 1% mark", angle= 90,
           color= "blue", size= 3)
```

The plot was then divided into geopolitical zones to compare states in the same zones
```{r}
p+ geom_point(mapping = aes(color=Zone))+ scale_x_log10(labels= scales::percent)+
  labs(x= "Percentage", y= "",
       title = "Percentage contribution of Nigeria states to the Country's IGR",
       subtitle = "First Quater of 2021",
       caption = "Data Source: National Bureau of Statistic",
       color= "Geopolitical Zones")+ facet_wrap(~Zone, scale= 'free_y')+ 
  guides(color= "none")
```

This finding showed that South west and the South south Zone of the country as the leading zones in term of IGR contribution of the country in the first quater of 2021, with North east zone having the lowest contribution.
Lagos, Rivers and FCT are the leading states while taraba, Niger and Yobe has the lowest contribution in that order.
